---
title: "Aggregating individuals to plot-mean value"
author: "Yunke Peng"
date: "Jan 14, 2022"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

### Preprocessing and select a specific plot as an example to show (duke_c, npp).
```{r}
library(dplyr)
library(metafor)
library(MAd)
library(tidyverse)
df <- read_csv("~/data/gcme/data_received_190325/NewData_wide_CORRECTED2.csv") %>%
  
  # something is wrong in columns ambient_Sd, ambient_Se, elevated...
  mutate( ambient_Sd  = as.numeric(ambient_Sd),  ambient_Se  = as.numeric(ambient_Se), 
          elevated_Sd = as.numeric(elevated_Sd), elevated_Se = as.numeric(elevated_Se) )

# save experiments names
df_experiments <- df %>% select(exp_nam, prev_name) %>% distinct()

# take only experiments >1 yr
list_exp_gt1yr <- df %>% 
  filter(!is.na(Year)) %>% 
  group_by(exp_nam) %>% 
  summarise(nyears=max(Year)) %>% 
  filter(nyears>1) %>% 
  select(exp_nam) %>% 
  unlist() %>% 
  unname()

df_c <- df %>%
  
  filter(treatment=="c") %>% 
  
  ## More than 1 year data
  filter(exp_nam %in% list_exp_gt1yr) %>% 
  
  ## Combine with experiments meta info
  left_join( 
    read_csv("~/data/gcme/data_received_190325/table_var_exp_names_experiments.csv") %>% 
      select(prev_name, Fumigation_type=my_fumigation_type, Vegetation_type),
    by = c("prev_name")
  ) %>% 
  
  ## Filter only Fumigation_type OTC or FACE
  filter( Fumigation_type %in% c("OTC", "FACE") ) %>%
  
  {.}
selvars <- c("NPP")

df_c_sub <- df_c %>% 
  filter(Data_type %in% selvars) %>% 
  mutate(varnam = "npp")

#calculating response ratio
df_c_sub <- df_c_sub %>%         
  
  # get standard deviation for all data
  mutate( 
    my_ambient_sd = ambient_Sd, 
    my_elevated_sd = elevated_Sd 
  ) %>%
  rowwise() %>% 
  mutate( 
    my_ambient_sd = ifelse( is.na(my_ambient_sd),  ambient_Se  * sqrt(n_plots), my_ambient_sd ),
    my_elevated_sd  = ifelse( is.na(my_elevated_sd), elevated_Se * sqrt(n_plots), my_elevated_sd )
  ) %>%
  
  ## Get logarithm of response ratio and its variance
  metafor::escalc( 
    measure = "ROM", 
    m1i = elevated, sd1i = my_elevated_sd, n1i = n_plots, 
    m2i = ambient,  sd2i = my_ambient_sd,  n2i = n_plots, 
    data=., 
    append = TRUE, var.names = c("logr", "logr_var") ) %>% 
  as_tibble() %>% 
  mutate( logr_se = sqrt(logr_var)/sqrt(n_plots) )

df_c_test <- subset(df_c_sub,prev_name=="DUKE_c" & varnam=="npp")

```

### Method 1: using MAd::agg function (the one previously used by Beni). But it has problem as discussed before, since the output value is exactly the same to normal individuals average.
```{r}

#method 1: using aggregation method 1: MAd::agg
df_plotmean1 <- df_c_test %>% 
  filter(!is.na(logr_var) & !is.na(logr)) %>% 
  mutate( id = paste(exp_nam, varnam, sep = "_XXX_")) %>% 
  MAd::agg( id = id, es = logr, var = logr_var, n.1 = n_plots, n.2 = n_plots, cor = 1.0, method = "BHHR", data = . ) %>% 
  as_tibble() %>% 
  mutate( id = str_split(id, "_XXX_") ) %>% 
  mutate( exp_nam = purrr::map_chr(id, 1),
          varnam = purrr::map_chr(id, 2) ) %>% 
  select(exp_nam, varnam, es, var)

df_plotmean1

#however, as discussed before (this function MAd::agg did have some problem), because it yield exactly the same value to normal mean values!
normal_mean <- df_c_test %>% 
  filter(!is.na(logr_var) & !is.na(logr)) %>%  group_by(exp_nam) %>%
  dplyr::summarize(Mean = mean(logr, na.rm=TRUE))

normal_mean

df_plotmean1$es - normal_mean$Mean

```


### Method 2: using metafor::rma.mv, as discussed and confirmed by Beni and Yunke in November (at UK). And we dicussed in one meeting that take measurement year as ramdon factor - did you remember it..? Now in all my analysis I have used this method as we confirmed earlier.
```{r}

#alternatively - we are using this method as previously confirmed from Beni - where random factor is year.
#method 2: using aggregation method 2: metafor
out_meta <- df_c_test %>% 
  filter(!is.na(logr_var) & !is.na(logr)) %>% 
  mutate( id = paste(exp_nam, varnam, sep = "_XXX_")) %>% 
  metafor::rma.mv( logr, logr_var, method = "REML",random = ~ 1 | Year, slab = Year,control = list(stepadj=0.3),data = .)

df_plotmean2 <- tibble(
  exp="DUKE_c",varnam="npp", es = out_meta$b[1,1],var= (out_meta$se*sqrt(out_meta$k))^2) # variance = (SD)^2 = (SE * sqrt(n_samples))^2


df_plotmean2
 

```
