---
title: "From multiple samples to plot-mean: two possible aggregation method?"
author: "Yunke Peng"
date: "Jan 14, 2022"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

### Preprocessing
reprocessing and select one specific plot and variable as an example to show (duke_c, npp).
```{r}
library(dplyr)
library(metafor)
library(MAd)
library(tidyverse)
df <- read_csv("~/data/gcme/data_received_190325/NewData_wide_CORRECTED2.csv") %>%
    mutate( ambient_Sd  = as.numeric(ambient_Sd),  ambient_Se  = as.numeric(ambient_Se), 
          elevated_Sd = as.numeric(elevated_Sd), elevated_Se = as.numeric(elevated_Se) )

# save experiments names
df_experiments <- df %>% select(exp_nam, prev_name) %>% distinct()

# take only experiments >1 yr
list_exp_gt1yr <- df %>% filter(!is.na(Year)) %>%  group_by(exp_nam) %>% summarise(nyears=max(Year)) %>% 
  filter(nyears>1) %>% select(exp_nam) %>% unlist() %>% unname()

df_c <- df %>%
  filter(treatment=="c") %>% 
  ## More than 1 year data
  filter(exp_nam %in% list_exp_gt1yr) %>% 
  ## Combine with experiments meta info
  left_join( 
    read_csv("~/data/gcme/data_received_190325/table_var_exp_names_experiments.csv") %>% 
      select(prev_name, Fumigation_type=my_fumigation_type, Vegetation_type),
    by = c("prev_name")
  ) %>% 
  ## Filter only Fumigation_type OTC or FACE
  filter( Fumigation_type %in% c("OTC", "FACE") ) %>%
  {.}

selvars <- c("NPP")

df_c_sub <- df_c %>% 
  filter(Data_type %in% selvars) %>% 
  mutate(varnam = "npp")

#calculating response ratio
df_c_sub <- df_c_sub %>%         
  # get standard deviation for all data
  mutate( 
    my_ambient_sd = ambient_Sd, 
    my_elevated_sd = elevated_Sd 
  ) %>%
  rowwise() %>% 
  mutate( 
    my_ambient_sd = ifelse( is.na(my_ambient_sd),  ambient_Se  * sqrt(n_plots), my_ambient_sd ),
    my_elevated_sd  = ifelse( is.na(my_elevated_sd), elevated_Se * sqrt(n_plots), my_elevated_sd )
  ) %>%
  ## Get logarithm of response ratio and its variance
  metafor::escalc( 
    measure = "ROM", 
    m1i = elevated, sd1i = my_elevated_sd, n1i = n_plots, 
    m2i = ambient,  sd2i = my_ambient_sd,  n2i = n_plots, 
    data=., 
    append = TRUE, var.names = c("logr", "logr_var") ) %>% 
  as_tibble() %>% 
  mutate( logr_se = sqrt(logr_var)/sqrt(n_plots) )

df_c_test <- subset(df_c_sub,prev_name=="DUKE_c" & varnam=="npp")

#response ratio, its variance, years of measurement, number of plots, and experiment name (DUKE_c)
summary(df_c_test[,c("logr","logr_var","Year","n_plots","prev_name")])

```

### Method 1: Mad::agg

using MAd::agg function (the one previously used). But it has problem, since the output value is exactly the same to normal individuals average.
```{r}

#method 1: using aggregation method 1: MAd::agg
df_plotmean1 <- df_c_test %>% 
  filter(!is.na(logr_var) & !is.na(logr)) %>% 
  mutate( id = paste(exp_nam, varnam, sep = "_XXX_")) %>% 
  MAd::agg( id = id, es = logr, var = logr_var, n.1 = n_plots, n.2 = n_plots, cor = 1.0, method = "BHHR", data = . ) %>% 
  as_tibble() %>% 
  mutate( id = str_split(id, "_XXX_") ) %>% 
  mutate( exp_nam = purrr::map_chr(id, 1),
          varnam = purrr::map_chr(id, 2) ) %>% 
  select(exp_nam, varnam, es, var)

df_plotmean1

#however,  it yield exactly the same value to mean values.
normal_mean <- df_c_test %>% 
  filter(!is.na(logr_var) & !is.na(logr)) %>%  group_by(exp_nam) %>%
  dplyr::summarize(Mean = mean(logr, na.rm=TRUE))

normal_mean

df_plotmean1$es - normal_mean$Mean

```


### Method 2: metafor::rma.mv
This method imputed variance of each sample (within one plot), and also considered  Years of treatment as random factor (or not). And then, the output "plot-mean" value is indeed different to above normal mean value, that is, necessarily affected by variance of each sample.

```{r}

#alternatively - we are using this method as previously confirmed from Beni - where random factor is year.
#method 2: using aggregation method 2: metafor
out_meta <- df_c_test %>% 
  filter(!is.na(logr_var) & !is.na(logr)) %>% 
  mutate( id = paste(exp_nam, varnam, sep = "_XXX_")) %>% 
  metafor::rma.mv( logr, logr_var, method = "REML",random = ~ 1 | Year, slab = Year,control = list(stepadj=0.3),data = .)

df_plotmean2 <- tibble(
  exp="DUKE_c",varnam="npp", es = out_meta$b[1,1],var= (out_meta$se*sqrt(out_meta$k))^2) # variance = (SD)^2 = (SE * sqrt(n_samples))^2

df_plotmean2
 
#if not considered measurement year as random factor..the output value differ slightly

out_meta <- df_c_test %>% 
  filter(!is.na(logr_var) & !is.na(logr)) %>% 
  mutate( id = paste(exp_nam, varnam, sep = "_XXX_")) %>% 
  metafor::rma.mv( logr, logr_var, method = "REML",control = list(stepadj=0.3),data = .)

df_plotmean3 <- tibble(
  exp="DUKE_c",varnam="npp", es = out_meta$b[1,1],var= (out_meta$se*sqrt(out_meta$k))^2) # variance = (SD)^2 = (SE * sqrt(n_samples))^2

df_plotmean3

```

So, we are wondering which method might best work if we are interested in plot-mean based meta-analysis?